Revisiting Multi-Domain Machine Translation

### 2020-04-17 2020/coling2020/multidomain.tex
start working on related work

We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Zhang16topicinformed}, register, style \cite{Niu18multitask}, etc. Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$. $h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$, but can differ from one domain to the other.

@@

This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is \cite{Farajian17multidomain}, where test translations rely on a similarity score between sentences, rather than domain labels.}, and one that we also adopt here. Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function. We now discuss the arguments that are usually put forward to proceed differently.
% --

@@

\section{Challenging multi-domain translation systems \label{sec:challenging}}
In this section, we discuss seven\fyTodo{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met.\fyTodo{Insert discussion about scores}

@@

In this study, all systems fail to outperform fine-tuning, sometimes by a wide margin: this is observed for an ``isolated'' domain such as \domain{rel}. This might be expected, since all domains are clearly separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other. In this situation, MDMT systems should make sure that each domain is allocated a sufficient number of specific parameters, so as to close the gap with fine-tuning. 
\fyDone{More comments when we have all the results}.\fyDone{Importance of sharing and unsharing}

\begin{table*}
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}|r|} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{news}} \\ \hline
    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 & 23.5\\
    \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 &\\
    \system{FT-Full}       & 37.7 & \bf 59.2 & \bf 54.5 & 34.0 & \bf 46.8 & \bf 90.8   & 42.8 & 53.8 &\\
   \system{FT-Block}     & 37.3 & \bf 57.9 & 53.9 & 33.8 & \bf 46.7 & \bf 90.2    & 42.3 & 53.3 &\\ \hline
 %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \bf 80.5 &41.7 & 50.1     & 21.8 \\
    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \bf 79.9 &41.4 & 49.9   & \sl 21.7 \\
    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \bf 79.9 &41.1 & 49.7       & \\
    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \bf 79.9 & & 49.8 & 22.1\\ 
    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 & \\
    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 & \\
    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 & \\
    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.1 & 51.9  & \SW{21.2} \\
    \hfill MDL Res (gen)           & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7&\\
%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3&\\
     \hline
  \end{tabular}
  \caption{Translation performance of various MDMT systems. We report the BLEU score for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages. Boldface denotes significant differences with respect to the \system{Mix-Nat} system. Scores for the \domain{news} domain correspond to the out-of-domain scenario (see Section~\ref{ssec:redomains}).}
  \label{tab:performance}
\end{table*}
\fyTodo{Fill the table with all results, including a comparison with *RNNs* for wcmd}\fyTodo{Do we have significancy tests for averages?}
\fyDone{A new column for the News domain with predicted domains (except for the generic case).}

@@

Table~\ref{tab:redomains} summaries the results of four experiments where we artificially redefine the limits of domains. In the three \textsl{split} experiments, we randomly split a domain corpus in two part: a good MDMT system should detect that these two ``domains'' should be mutually beneficial, and should be hardly affected by this change wrt.\ the baseline scenario. This is the situation where we should see clear improvements even over finetuning, which only exploits part of the available relevant data. In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains. \fyTodo{Prepare table for the wrong domain analysis (see also last col. of the first table)}In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.

\begin{table}
  \centering
  \begin{tabular}{|p{2.5cm}|*{9}{r|}} \hline
    % &&&&&& \\
    \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & Wrong \\ % \hline
     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{bank} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} & domain \\ \hline
    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law} & \\
    \system{FT-Full}      & -0.13 & -0.6 & -1.51& -0.25& & & & & \\
    \system{DC-Tag}     & -0.16 & -0.27& +0.13& +0.2& & & & & \\
    \system{DC-Feat}    & -0.47& 0.0 & +0.27 $^+$ & +0.27 $^+$& & & & & \\
    \system{TTM}          & & & & & & & & &  \\
    \system{LDC-Tag}   & & & & & & & & & \\

    \system{MDL Res}   & -0.2$^+$& -0.07$^+$&+0.18$^+$  &+0.0$^+$ & & & & & \\
%    \system{MDL Gated} & & & & & & & & & \\

    \system{ADM} & & & & & & & & & \\
    \system{DM}    & & & & & & & & & \\
    \system{WDC}  & & & & & & & & & \\
    \hline
  \end{tabular}
  \caption{Translation performance with variable domain definitions. For the Split experiments, we only report the loss in performance for the \domain{med} test set, (*) denote significant losses when domains are split or merged; (+) denotes a significant improvement over \system{Full FT}; \fyTodo{How about wrong ?}}
  \label{tab:redomains}
\end{table}

@@

\subsubsection{Handling unknown domains \label{sssec:unknowns}}
To evaluate requirement [P5-UNK], we translate sentences from a domain unseen in training (\domain{news}). For each test sentence, we automatically predict the most likely domain tag, and use it for decoding. Results are in the rightmost column of Table~\ref{tab:performance}.

@@

The multi-domain training regime is more the norm than the exception for NLP \cite{Dredze08online}. We focus here mostly on multi-domain learning in machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection and weighting, adversarial training, etc) have also been proposed for other language processing tasks.
%; scores are reported separately per domain, suggesting that all domains are equally important at test time.
% Adversarial adpatation is used in many domains
% for QA @misc{lee2019domainagnostic,
%    title={Domain-agnostic Question-Answering with Adversarial Training},
%    author={Seanie Lee and Donggyu Kim and Jangwon Park},
%    year={2019},
%    eprint={1910.09342},
%    archivePrefix={arXiv},
%    primaryClass={cs.CL}
%}, NLU and many others

The authors of \cite{Kobus17domaincontrol} propose to inject additional domain feature in the model, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word. This technique has also been used to control the style of MT outputs in \cite{Sennrich16politeness}, and to the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}. This approach is extended in \cite{Pham19generic} where source word embeddings are likely to vary across domains: this is achieved by spliting the embedding vector into a generic and a domain-specific part (one for each possible domain). Such feature-based approaches strongly rely on earlier proposals from the era of statistical MT \cite{Daume07frustratingly,Clark12onesystem}. Domain control can also be included in the target side, as in \cite{Chen16guided}, where a vector of topic describing the current document is included as an extra context in the softmax layer of the decoder.

>>>>

We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc. Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$. $h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$, but can differ from one domain to the other.

@@

This definition seems to be the most accepted view of a multi-domain MT\footnote{One important exception is \cite{Farajian17multidomain}, where test translations rely on a similarity scores between test and train sentences, rather than on domain labels.}, and one that we also adopt here. Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function. We now discuss the arguments that are usually put forward to proceed differently.
% --

@@

\section{Challenging multi-domain translation systems \label{sec:challenging}}
In this section, we discuss seven\fyTodo{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met. All these evaluations will rest on comparison of translation performance, and do not depend on the choice a particular metrics. To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best choice to measure subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}. Other important figure of merits for MDMT systems are the computational cost and the total number of parameters.\fyDone{Insert discussion about scores}

@@

In this study, all systems fail to outperform fine-tuning, sometimes by a wide margin: this is observed for an ``isolated'' domain such as \domain{rel}. This might be expected, since all domains are clearly separated (cf.\ Section~\ref{ssec:corpora}) and are hardly helping each other. In this situation, MDMT systems should make sure that each domain is allocated a sufficient number of specific parameters, so as to close the gap with fine-tuning.
\fyDone{More comments when we have all the results}\fyDone{Importance of sharing and unsharing}

\begin{table*}
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}|r|} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} & \multicolumn{1}{c|}{\domain{news}} \\ \hline
    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 & 23.5\\
    \system{Mixed-Bal}   &  35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5 & 40.3  & 51.4 &\\
    \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8 &\\
   \system{FT-Block}     & 37.3 & \SB{57.9} & 53.9 & 33.8 & \SB{46.7} & \SB{90.2}  & 42.3 & 53.3 &\\ \hline
 %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & \SB{80.5} &41.7 & 50.1     & 21.8 \\
    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & \SB{79.9} &41.4 & 49.9   & \SW{21.7} \\
    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & \SB{79.9} &41.1 & 49.7       & \\
    \system{LDC-Tag}     & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & \SB{79.9} & & 49.8 & 22.1\\ 
    \system{DM}             & \SW{35.6} & \SW{49.5}  & \SW{45.6}& \SW{29.9} & \SW{37.1} & \SW{62.4} & 39.7 & 43.8 & \\
    \system{ADM}           & 36.4 & \SW{53.5}  & \SW{48.3} & \SW{32.0} & \SW{41.5} & \SW{73.4} & 40.7 & 47.5 & \\
    \system{WDCMT}       & 34.6 & 51.7 & 47.5 & 29.4 & 36.5 & 49.3 & & 41.5 & \\
    \system{MDL Res}     & 37.9 & \SB{56.0}  & \SB{51.2}   & 33.5   &  44.4  & \SB{88.3} & 42.1 & 51.9  & \SW{21.2} \\
    \hfill MDL Res (gen)    & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7&\\
%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3&\\
     \hline
  \end{tabular}
  \caption{Translation performance of various MDMT systems. We report the BLEU score for each domain, as well as domain-weighted (w\domain{avg}) and unweighted (\domain{avg}) averages. Boldface denotes significant differences with respect to the \system{Mix-Nat} system. Scores for the \domain{news} domain correspond to the out-of-domain scenario (see Section~\ref{ssec:redomains}).}
  \label{tab:performance}
\end{table*}
\fyTodo{Fill the table with all results, including a comparison with *RNNs* for wcmd}\fyTodo{Do we have significancy tests for averages?}
\fyDone{A new column for the News domain with predicted domains (except for the generic case).}

@@

Table~\ref{tab:redomains} summaries the results of four experiments where we artificially redefine the limits of domains, with the aim to challenge requirements [P2-TUN], [P3-HET], and [P4-ERR]. In the three \emph{split} experiments, we randomly split a domain corpus in two part: a good MDMT system should detect that these two ``domains'' should be mutually beneficial, and should be hardly affected by this change wrt.\ the baseline scenario. This is the situation where we should expect to MDMT even surpass fine-tuning separately on each of these dummy domain, as MDMT exploits all data, where fine-tuning only sees a subpart.

In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains. \fyTodo{Prepare table for the wrong domain analysis (see also last col. of the first table)}In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.

\begin{table}
  \centering
  \begin{tabular}{|p{2.5cm}|*{9}{r|}} \hline
    % &&&&&& \\
    \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & Wrong \\ % \hline
     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{bank} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} & domain \\ \hline
    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law} & \\
    \system{FT-Full}      & -0.13 & -0.6 & -1.51& -0.25& & &\SW{-1.6} & \SW{-1.4}& \\
    \system{DC-Tag}     & -0.16 & -0.27& +0.13& +0.2& & & -0.5 & -0.4 & \\
    \system{DC-Feat}    & -0.47& 0.0 & \SB{0.27} $^+$ & +0.27 $^+$& & & 0.3 & 0.1 & \\
    \system{TTM}          & & & & & & &  0.0 & -0.3 &  \\
    \system{LDC-Tag}   & & & & & & &  0.0 & 0.1& \\ 
    \system{MDL Res}   & -0.2$^+$& -0.07$^+$&+0.18$^+$  &+0.0$^+$ & & & \SW{-2.4} & \SW{-3.5} & \\
%    \system{MDL Gated} & & & & & & & & &  \\
    \system{ADM} & & & & & & &  0.1 & -0.4 & \\
    \system{DM}    & & & & & & & 0.9 & 0.1 & \\
    \system{WDC}  & & & & & & & & & \\
    \hline
  \end{tabular}
  \caption{Translation performance with variable domain definitions. For the Split experiments, we only report the loss in performance for the \domain{med} test set, (*) denote significant losses when domains are split or merged; (+) denotes a significant improvement over \system{Full FT}; \fyTodo{How about wrong ?}}
  \label{tab:redomains}
\end{table}

@@

It is finally interesting to note that all systems, but \system{MDL Res}, are quite robust to domain merge: we see even small gains in performance (eg.\ for \system{DM}). While robustness is generally a good thing, it is also a sign of a poor overall usage of domain labels. \system{MDL Res}\fyTodo{To be completed}

\subsubsection{Handling unknown domains \label{sssec:unknowns}}
To evaluate requirement [P5-UNK], we translate sentences from a domain unseen in training (\domain{news}). For each test sentence, we automatically predict the most likely domain tag, and use it for decoding. Results are in the rightmost column of Table~\ref{tab:performance}.\fyTodo{To be completed}

@@

The multi-domain training regime is more the norm than the exception for NLP \cite{Dredze08online,Finkel09hierarchical}. We focus here mostly on multi-domain learning in machine translation, keeping in mind that similar problems and solutions (parameter sharing, instance selection and weighting, adversarial training, etc) have also been proposed for most language processing tasks.
%; scores are reported separately per domain, suggesting that all domains are equally important at test time.
% Adversarial adpatation is used in many domains
% for QA @misc{lee2019domainagnostic,
%    title={Domain-agnostic Question-Answering with Adversarial Training},
%    author={Seanie Lee and Donggyu Kim and Jangwon Park},
%    year={2019},
%    eprint={1910.09342},
%    archivePrefix={arXiv},
%    primaryClass={cs.CL}
%}, NLU and many others

Multi-domain translation was introduced for statistical MT (eg.\ \cite{Banerjee10combining,Sennrich13multidomain}): as for domain adaptation, two main strategies where considered consisting at weighting train and test instances based on their domain-relatedness on the one hand; or to inject domain labels as additional features in learning.

This latter strategy has also been widely used in MNT: the authors of \cite{Kobus17domaincontrol} inject an additional domain feature in their seq2seq model, either in the form of an extra (initial) domain-token, or in the form of additional domain-feature associated to each word. Note his technique has also been used to control the style of MT outputs in \cite{Sennrich16politeness,Niu18multitask}, and to encode the source or target languages in multilingual MT \cite{Firat16multiway,Johnson17google}. This approach is extended in \cite{Pham19generic} where source word embeddings are likely to vary across domains: this is achieved by splitting the embedding vector into a generic and a domain-specific part (one for each possible domain); a similar approach, applied to hidden states, is documented in \cite{Chu18multilingual}. Such feature-based approaches strongly rely on earlier proposals from the era of statistical MT \cite{Daume07frustratingly,Clark12onesystem}. Domain control can also be performed on the target side, as in \cite{Chen16guided}, where a vector of topic describing the current document is included as an extra context in the softmax layer of the decoder.
